---
title: "MA [46]15 Homework 1"
author: "Maika Takemoto"
format: gfm
---

In this homework I'm analyzing a small subset of the IPUMS dataset.
I start by loading the packages and the data.

```{r init}
library(tidyverse)
library(ggthemes)
load("ipums_hw_data.RData")
ipums_hw_data
```

Descriptions for the variables in the data set are given below.

| Variable | Description       |
|:--------|:-------------------|
|YEAR     | reports the four-digit year when the household was enumerated or included in the census, the ACS, and the PRCS. |
|SERIAL   | An identifying number unique to each household record in a given sample. |
|HHINCOME | Reports the total money income of all household members age 15+ during the previous year. |
|PERWT    | Indicates how many persons in the U.S. population are represented by a given person in an IPUMS sample. |
|RACE     | Simple version of race. The concept of race has changed over the more than 150 years represented in IPUMS. Currently, the Census Bureau and others consider race to be a sociopolitical construct, not a scientific or anthropological one. Many detailed categories consist of national origin groups. |
|RACED    | Detailed version of race.  |
|POVERTY  | Expresses each family's total income for the previous year as a percentage of the poverty thresholds established by the Social Security Administration. |
|SPMPOV   | Indicates whether a family is classified as poor according to the Supplemental Poverty Measure (SPM). |

The number of observations in the data set is less than 0.01% of the full IPUMS data for the relevant years.
This data was collected from [IPUMS USA](https://usa.ipums.org/) and was cleaned and reduced in size to improve ease-of-use.
You can search for individual variables [here](https://usa.ipums.org/usa-action/variables/search).


## Question 1
From the graph, it can be observed that household income and poverty levels have a positive correlation. A significant cluster of points appears around higher income levels, but with varying poverty levels. This suggests that even at higher income levels, some populations experience degrees of poverty. There also arises a plateau in poverty levels around log(HHINCOME) of 12.5, indicating that those at the highest income level rarely experience poverty.


```{r q1}
ggplot(ipums_hw_data) + 
  geom_point(aes(x = log(HHINCOME), y = log(POVERTY), color = RACE), alpha = 0.25) +
  scale_color_colorblind() + 
  labs(title = "Log-Log Plot of Poverty vs. Household Income", x = "HHINCOME", y = "POVERTY") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Question 2

```{r q2}
ggplot(ipums_hw_data,aes(x = HHINCOME, y = RACE, fill = RACE)) +
  geom_boxplot(varwidth = TRUE) +
  scale_fill_colorblind(guide = guide_none()) +
  scale_x_discrete() 
```

## Question 3

```{r q3}
ggplot(ipums_hw_data, aes(x = RACE, fill = SPMPOV, weight = PERWT)) + 
  geom_bar(position = "fill") + 
  scale_x_discrete(guide = guide_axis(angle = 15)) +  
  scale_fill_discrete(name = NULL) + 
  labs(y = "Proportion") +  
  scale_y_continuous(labels = scales::percent)  
```
Note - `position = "fill"` in `geom_bar` call and `weight = PERWT` in `aes` call make the bars proportional to each other by summing each bar to equal 100%. The bars are then weighted by population to ensure that the visualized proportions are more accurate.

## Question 4

```{r q4}
knitr::include_graphics("raced_vs_mean_hhincome.svg")

ggplot(ipums_hw_data,aes(x = HHINCOME, y = RACE)) + 
  stat_summary(aes(x = HHINCOME, y = RACE, color = RACED), fun = mean, size = 0.20) + 
  stat_summary(aes(x = HHINCOME, y = RACE), fun = mean, geom = "point", size = 4, alpha = 0.4) +
  theme(legend.position = "none") +
  ggtitle("Mean household income by detailed race", subtitle = "Large grey dots are means for simplified race variable") +
  scale_y_discrete(name = " ") +
  scale_x_continuous(name = "mean household income", labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) 
```

Q. What are some important differences between the figures? 

One key difference between the figures is that in the homework, only the mean household income is graphed. In the document, it presents the poverty rates within the various races. Additionally, the latter figure also presents us with a wider distribution of the poverty rates compared to our plotting distribution of the average household incomes.

Q. Are there any issues with the figure presented in this homework?

In the figure in the homework, it is unclear what "Three or more major races" and "Two major races" implies. Also, for the data concerning "Japanese", Black/African American", and "White", the tiny points that represent the mean of HHINCOME for each RACE group are the same as the mean for the simplified RACE variable, giving for a not so effective visualization of the data.

