---
title: "MA [46]15 Homework 5"
author: "Maika Takemoto"
format: gfm
---


## Question 1


In this work I'll analyze the titles of the most recent 10,000 papers in the
[Criminal Justice category](https://www.ssrn.com/index.cfm/en/cjrn/) of the
[Social Science Research Network](https://www.ssrn.com/index.cfm/en/) preprint
repository. The data in file `data/ssrn_cjn.rds` is loaded into the table
`papers`. As an initial check, I'm plotting the distribution of number of words
in title for each year.

```{r q1}
suppressPackageStartupMessages(library(tidyverse))

# Read in the data
papers <- readRDS("data/ssrn_cjn.rds")

# Create a new variable `word_count`
papers <- papers |>
  mutate(
    word_count = str_count(title, "\\b\\w+\\b")
  )

# Plot the boxplot 
ggplot(papers, aes(x = as.factor(year), y = word_count)) +
  geom_boxplot(notch = TRUE) +
  coord_flip() +
  labs(
    title = "Word Count in Titles per Year",
    x = "Year",
    y = "Word Count"
  )

# Calculate the number of titles with exactly one word
one_word_titles <- papers |>
  filter(word_count == 1) |>
  nrow()

one_word_titles
```

The box plot above shows an overview of the distribution of various titles' word counts across the past 5 years. The interquartile ranges across the years appear fairly consistent, with the majority of word counts lying within 25 - 50. There is also a noticeable stability in the median word counts across the years. However, each year, especially 2021, has several outliers on the higher end.  

There are 17 papers with one-word titles, which likely appear as outlieras on the lower end of each year's data points.

## Question 2

```{r q2}
keywords <- tibble(keyword = 
                     c("crimin", "justice", "polic", 
                       "eviden", "violen", "right", "international")
                   )
# example of crossing between two tables:
# crossing(tibble(x = 1:3), tibble(y = letters[1:3]))

papers_keywords <- papers |>
  crossing(keywords) |>
  mutate(
    keyword_in_title = str_detect(str_to_lower(title), keyword)
  ) |>
  count(year, keyword, wt = keyword_in_title * downloads)

# Reshape data 
papers_summary <- papers_keywords |>
  pivot_wider(
    names_from = keyword,
    values_from = n,
    values_fill = 0
  )

papers_summary
```


## Question 3

```{r q3}
papers_summary |>
  pivot_longer(
    cols = -year,  
    names_to = "keyword",
    values_to = "n"
  ) |>
  ggplot(aes(x = year, y = n)) +
  geom_line(aes(color = keyword), size = 1) +
  labs(
    title = "Download Counts for Keywords in Paper Titles",
    x = "Year",
    y = "Download Counts",
    color = "Keyword"
  ) +
  theme_minimal()
```

The most commonly downloaded keyword is "crimin", which spiked in 2019. In 2020, there was a spike for both "violen", "eviden", and "polic", which may have been in correlation with Covid-19. However, after 2020, there was an overall decrease for all of the keywords with an all time low reached in 2023. 

## Question 4

```{r q4}
# Calculate total downloads per year
papers_year <- papers |>
  group_by(year) |>
  summarize(total_downloads = sum(downloads, na.rm = TRUE))

# Join the total downloads with the keyword counts dataset
papers_prop <- papers_keywords |>
  left_join(papers_year, by = "year") |>
  mutate(prop = n / total_downloads)  

# Plot the proportions
papers_prop |>
  ggplot(aes(x = year, y = prop, color = keyword)) +
  geom_line(size = 1) +
  labs(
    title = "Total Downloads per Keyword by Year",
    x = "Year",
    y = "Total Downloads",
    color = "Keyword"
  ) +
  theme_minimal()
```

With this proportional plot, there is more variability to be seen in the trend for the keyword "crimin". It also had a spike in 2022 rather than 2019. However, the keyword "violen" and "polic" had the same spike in 2020 with the former following the same trend in both graphs. The least downloaded word, "international", was the same for both plots as well.
